# ---
# title: "Phonotactic Network Generator"
# author: Leo Niehorster-Cook
# output: html_notebook
# ---
In this file, we will generate the full biphonemic network. 
On this network, A and B are neighbors iff they share a substring of length 2. 
The standard English word list has nearly 20,000 entrees, impeding computation.
For tests, we will use Toki Pona, with 120 words, so that everything is fast.
Note that Toki Pona is written phonetically, with a subset of English letters.
Thus, for Toki Pona, the orthography happens to be equivalent to the Klattese. 
However, this file will work for any lexicon.

To make the network, we're going to have to load up our data and our packages.
Then, we'll define some useful functions before building anything.
Before we build the biphonemic network, we'll build a bipartite network.
The layers will represent words and biphones. Edges will denote substrings.
The phonotactic network is then just the projection of that.
 
Let's load our packages and our data.

```{r} 
# Packages

library(magrittr)
library(data.table) 
library(igraph)
library(ggplot2)
library(spreadr)
```


```{r} 
# Data - Pick your language

#data <- 'pu.csv'
data <- 'WordList.csv' # English

lexicon.dt <- fread(file=data)

```

<!-- Now we've loaded everything up, so let's define some useful functions. -->

```{r} 
allsubstr <- # Returns an array of all unique substrings of length n from x 
  function(x, n) 
    substring(
      text = x,
      first = 1:(nchar(x) - n + 1),
      last = n:nchar(x)
    ) %>% 
    unique()

kgrams <- # Concatenates the substrings with a period to mark nonlexicality
  function(x, n) '.' %>% paste0(allsubstr(x = x, n = n)) 

biphones <- # Returns all the "2grams," i.e. two-length kgrams
  function(x) kgrams(x = x, n = 2)
```

<!-- Now biphones('aardvark') returns: -->
<!-- [1] ".ar" ".rd" ".dv" ".va" ".aa" ".rk" -->

<!-- Note that biphones are represented as strings that start with a period.  -->
<!-- This is to differentiate them from words, such that <to> != /to/. -->

<!-- Note further that ".ar" is only listed once. -->

```{r}

lexicon.dt[,constituents:=Vectorize(biphones)(Klattese)]

# Now our lexicon is stored as a data table
# That column contains the constituent biphones of each word. 
# Let's take a peek inside:

head(lexicon.dt)
```
<!-- Our lexicon now has a set of constituent biphones associated with each word.  -->

<!-- To build the bipartite network, we'll need the vertex set and edge set. -->
<!-- Every word is a vertex, as is every biphone, marked by an initial '.' -->
<!-- The edges will be between words and biphones that are their substrings. -->

```{r}
biphoneset <-
  lexicon.dt[,sort(unique(unlist(constituents)))]

head(biphoneset)
```
<!-- Now we have a set of all biphones, with '.' preceeding the string. -->
<!-- Let's construct a dataframe that holds the vertices, i.e. words and biphones. -->
<!-- In this dataframe, words will be named in orthographically, not Klattese. -->
<!-- Again, in Toki Pona, this is redundant. -->

<!-- We'll also mark each vertex for lexicality by "type". -->
<!-- The first column will be the wordlist followed by the set of biphones. -->
<!-- The second column will be 'TRUE' for words and 'FALSE' for biphones. -->
```{r}
vset <- data.frame(
  v = c(
    lexicon.dt$Orthography, 
    biphoneset
  ),
  type = c(
    rep(TRUE,length(lexicon.dt$Orthography)),
    rep(FALSE,length(biphoneset))
  ),
  length = c(
    lexicon.dt$Orthography %>% nchar(),
    rep('N/A',length(biphoneset))
  )
)

```

<!-- We're nearly ready to construct the graph. -->
<!-- Before we do, let's get the edges into another dataframe. -->
```{r}
eset <- lexicon.dt[,
  list(
    Orthography = rep(Orthography,lengths(constituents)),
    biphone = unlist(constituents)
  )
]
```

<!-- Now let's build it! -->
```{r}
parser.g <- graph_from_data_frame(d = eset, vertices = vset)
```

Great! Now we have the bipartite network, 'b'. While we're at it, let's generate
the lexical projection, 'w', and the biphonemic projection, 'b'.
```{r}
lexicon.g <- # The Lexicon
  bipartite_projection(
    graph = parser.g,
    which="true"
  )

sublexicon.g <- # The Sublexicon 
  bipartite_projection(
    parser.g,
    which="false"
  )
```


Let's write the networks to memory and get out of here!


```{r}
 write_graph(
   graph = lexicon.g,
   file = 'lexicon.gml',
   format = 'gml'
  )

 write_graph(
   graph = sublexicon.g,
   file = 'parser.gml',
   format = 'gml'
  )
    
```

