# ---
# title: "Lexical Network Analyzer"
# author: Leo Niehorster-Cook
# ---

As you may have guessed, this notebook is for analyzing lexical networks. Thus
far, it's fit to analyze networks generated by PhonotacticNetworkGenerator,
but hopefully with time it will become broadly suited for analyzing network
representations of linguistic items.
 
```{r} 
# Import Libraries

library(magrittr)
library(igraph)
library(spreadr)
library(data.table)

```

Since we'll be focusing on the lexicon, "lex," let's define some functions that
specify it so we can save some time.

```{r}
# This function takes in a graph, a vertex, and a length (n), 
# and returns the n-hop neighborhood of that vertex. Defaults to 1
nhophood <- 
  function(v, h = 1) lexicon.g %>%  
    induced_subgraph(
      vids = unlist(
        neighborhood(
          graph = lexicon.g,
          nodes = v,
          order = h
        )
      )
    )
```

```{r}
# This is a general "inspect" function; I happen to like these layouts.
# inspect <- 
#   function(network) network %>% 
#   plot(
#     layout = layout_with_graphopt(
#       network,
#       charge = 100, mass = .01,max.sa.movement = 2,
#       spring.length = 100000,spring.constant = .01
#     )
#   )

inspect <- 
  function(network) network %>% 
  plot(
    layout = layout.davidson.harel(lexicon.g)
    )

```

```{r}
# This function combines the last two. I prefer a different layout when I am
# specifically inspecting n-hop neighborhoods.

inspect_hood <- function(v, h = 2) nhophood(v, h) %>% 
  plot(
    layout = layout_as_tree(
    graph = nhophood(v, h),
    root = v,
    circular = TRUE
  )
)
```

```{r}
# This function returns the clustering coefficient, C(x), given just the word
# Note that igraph defines C(x) differently than as in Watts & Strogatz (1998)

cc <- 
  function(v) transitivity(graph = lexicon.g,type = "local",vids =v)

add_cc <-
  function(dt,word) data.table(dt)[,clustering:=cc(word)]

add_degree <- 
  function(dt,word) data.table(dt)[,degree:=degree(lexicon.g,word)]
```

```{r}

```

